---
title: "Model_Exploration"
output: html_document
date: "2023-12-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tseries)
library(ggplot2)
library(tidyverse)
library(reshape2)
library(car)
library(lubridate)
library(lme4)

```

Reading in the documents I need:
```{r}
brent_crude_prices<- read_excel("brent_crude_prices.xlsx")
co2_emissions_data<-read.csv("co2_emission.csv")
ftse<- read.csv("FTSE 100 Historical Data.csv")
eu_ets_data<-read_excel("EU-ETS-data.xlsx")
```

Converting dates to the right format so I can join them:

```{r}
eu_ets_data$Date<- as.Date(eu_ets_data$Date)
brent_crude_prices$Date<- as.Date(brent_crude_prices$Date)
ftse$Date<- mdy(ftse$Date)
ftse$Price<- as.numeric(str_replace(ftse$Price,",",""))
```

Joining the three data frames:

```{r}
auction_and_oil<- merge(x=eu_ets_data,y=brent_crude_prices, 
                         by="Date", all.x=TRUE) 

auction_and_oil<-auction_and_oil |> filter(is.na(auction_and_oil$`Europe Brent Spot Price FOB (Dollars per Barrel)`)==FALSE)

oil_prices<- auction_and_oil$`Europe Brent Spot Price FOB (Dollars per Barrel)`
pacf(oil_prices, lag=10, pl=TRUE)

ets_stock_and_oil_data<- merge(x=auction_and_oil, y= ftse, by='Date', all.x=TRUE)
```

Now we need to clean the ets_stock and oil data so we can get an appropriate lag and check autocorrelations for each type of data:
```{r}
#First let's change auction name to just have EU, DE, PL or EUAA:

ets_stock_and_oil_data<- ets_stock_and_oil_data |> mutate(
                AuctionType = ifelse(str_detect(`Auction Name`, "EUAA")==TRUE,
                                                                        "EUAA",
                                                    str_sub(`Auction Name`,-2,)
), .after=`Auction Name`)
```

Setting up lags, starting with EU_only_auction:
```{r}
EU_only_auction<- ets_stock_and_oil_data |> filter(AuctionType=="EU") |> filter(
                                          is.na(`Auction Price €/tCO2`)==FALSE
)
eu_only_auction_price<-EU_only_auction$`Auction Price €/tCO2`

pacf(eu_only_auction_price, lag=10, pl=TRUE)

#Auction Price Lag
Auction_Price_lag<- EU_only_auction$`Auction Price €/tCO2`[1:length(EU_only_auction$`Auction Price €/tCO2`)-1]
Auction_Price_lag<- c(0,Auction_Price_lag)
EU_only_with_lag<- EU_only_auction |> mutate(
                                          Auction_Price_lag= Auction_Price_lag, .after=`Auction Price €/tCO2`
)

#Maximum bid lag
Max_bid_lag<- EU_only_auction$`Maximum Bid €/tCO2`[1:length(EU_only_auction$`Maximum Bid €/tCO2`)-1]
Max_bid_lag<- c(0,Max_bid_lag)
EU_only_with_lag<- EU_only_with_lag |> mutate(
                                          Max_bid_lag= Max_bid_lag, .after=`Maximum Bid €/tCO2`
)

#Avg bids per bidder lag
Avg_num_bids_per_bidder_lag<- EU_only_auction$`Average number of bids per bidder`[1:length(EU_only_auction$`Average number of bids per bidder`)-1]
Avg_num_bids_per_bidder_lag<- c(0,Avg_num_bids_per_bidder_lag)
EU_only_with_lag<- EU_only_with_lag |> mutate(
                                          Avg_num_bids_per_bidder_lag= Avg_num_bids_per_bidder_lag, .after=`Average number of bids per bidder`
)

#Avg vol per bidder
Avg_vol_per_bidder_lag<- EU_only_auction$`Average volume bid per bidder`[1:length(EU_only_auction$`Average volume bid per bidder`)-1]
Avg_vol_per_bidder_lag<- c(0,Avg_vol_per_bidder_lag)
EU_only_with_lag<- EU_only_with_lag |> mutate(
                                          Avg_vol_per_bidder_lag= Avg_vol_per_bidder_lag, .after=`Average volume bid per bidder`
)

#oil price lag
EU_only_with_lag<- EU_only_with_lag |> rename(
                        oil_price = `Europe Brent Spot Price FOB (Dollars per Barrel)`
)
oil_price_lag<- EU_only_with_lag$oil_price[1:length(EU_only_with_lag$oil_price)-1]
oil_price_lag<- c(0,oil_price_lag)
EU_only_with_lag<- EU_only_with_lag |> mutate(
                                          oil_price_lag= oil_price_lag, .after= oil_price
)

#ftse price lag:
EU_only_with_lag<- EU_only_with_lag |> rename(
                        ftse_price = Price
)
ftse_price_lag<- EU_only_with_lag$ftse_price[1:length(EU_only_with_lag$ftse_price)-1]
ftse_price_lag<- c(0,ftse_price_lag)
EU_only_with_lag<- EU_only_with_lag |> mutate(
                                          ftse_price_lag= ftse_price_lag, .after= ftse_price
)

#Auction Volume tCO2 lag:

Auction_vol_lag<- EU_only_auction$`Auction Volume tCO2`[1:length(EU_only_auction$`Auction Volume tCO2`)-1]
Auction_vol_lag<- c(0,Auction_vol_lag)
EU_only_with_lag<- EU_only_with_lag |> mutate(
                                          Auction_vol_lag= Auction_vol_lag, .after=`Auction Volume tCO2`
)
##

EU_only_auction_variables_fit<- glm(`Auction Price €/tCO2`~ Auction_Price_lag+ `Auction Volume tCO2`+ `Maximum Bid €/tCO2`,
                            family=gaussian, data=EU_only_with_lag)

durbinWatsonTest(EU_only_auction_variables_fit)
summary(EU_only_auction_variables_fit)

#DW of 1.27. Not terrible. How do we push this up to acceptable levels?


#EU_only_oil_fit<- glm(`Auction Price €/tCO2`~ lag+ `Auction Volume tCO2`+ `Maximum Bid €/tCO2`+
 #                                   oil_price,
  #                          family=gaussian, data=EU_only_with_lag)
#summary(EU_only_oil_fit)
#durbinWatsonTest(EU_only_oil_fit)


#So the difference in DW witha nd without oil prices isn't huge. Only 0.01 difference in the test statistic. 

```

Lag setup for just the German auctions:

```{r}
DE_only_auction<-ets_stock_and_oil_data |> filter(AuctionType=="DE") |> filter(
                                          is.na(`Auction Price €/tCO2`)==FALSE
)
DE_only_auction_price<- DE_only_auction$`Auction Price €/tCO2`
pacf(DE_only_auction_price)


#Auction Price Lag
Auction_Price_lag<- DE_only_auction$`Auction Price €/tCO2`[1:length(DE_only_auction$`Auction Price €/tCO2`)-1]
Auction_Price_lag<- c(0,Auction_Price_lag)
DE_only_with_lag<- DE_only_auction |> mutate(
                                          Auction_Price_lag= Auction_Price_lag, .after=`Auction Price €/tCO2`
)

#Maximum bid lag
Max_bid_lag<- DE_only_auction$`Maximum Bid €/tCO2`[1:length(DE_only_auction$`Maximum Bid €/tCO2`)-1]
Max_bid_lag<- c(0,Max_bid_lag)
DE_only_with_lag<- DE_only_with_lag |> mutate(
                                          Max_bid_lag= Max_bid_lag, .after=`Maximum Bid €/tCO2`
)

#Avg bids per bidder lag
Avg_num_bids_per_bidder_lag<- DE_only_auction$`Average number of bids per bidder`[1:length(DE_only_auction$`Average number of bids per bidder`)-1]
Avg_num_bids_per_bidder_lag<- c(0,Avg_num_bids_per_bidder_lag)
DE_only_with_lag<- DE_only_with_lag |> mutate(
                                          Avg_num_bids_per_bidder_lag= Avg_num_bids_per_bidder_lag, .after=`Average number of bids per bidder`
)

#Avg vol per bidder
Avg_vol_per_bidder_lag<- DE_only_auction$`Average volume bid per bidder`[1:length(DE_only_auction$`Average volume bid per bidder`)-1]
Avg_vol_per_bidder_lag<- c(0,Avg_vol_per_bidder_lag)
DE_only_with_lag<- DE_only_with_lag |> mutate(
                                          Avg_vol_per_bidder_lag= Avg_vol_per_bidder_lag, .after=`Average volume bid per bidder`
)

#oil price lag
DE_only_with_lag<- DE_only_with_lag |> rename(
                        oil_price = `Europe Brent Spot Price FOB (Dollars per Barrel)`
)
oil_price_lag<- DE_only_with_lag$oil_price[1:length(DE_only_with_lag$oil_price)-1]
oil_price_lag<- c(0,oil_price_lag)
DE_only_with_lag<- DE_only_with_lag |> mutate(
                                          oil_price_lag= oil_price_lag, .after= oil_price
)

#ftse price lag:
DE_only_with_lag<- DE_only_with_lag |> rename(
                        ftse_price = Price
)
ftse_price_lag<- DE_only_with_lag$ftse_price[1:length(DE_only_with_lag$ftse_price)-1]
ftse_price_lag<- c(0,ftse_price_lag)
DE_only_with_lag<- DE_only_with_lag |> mutate(
                                          ftse_price_lag= ftse_price_lag, .after= ftse_price
)

#Auction Volume tCO2 lag:

Auction_vol_lag<- DE_only_auction$`Auction Volume tCO2`[1:length(DE_only_auction$`Auction Volume tCO2`)-1]
Auction_vol_lag<- c(0,Auction_vol_lag)
DE_only_with_lag<- DE_only_with_lag |> mutate(
                                          Auction_vol_lag= Auction_vol_lag, .after=`Auction Volume tCO2`
)

#DE_only_oil_fit<- glm(`Auction Price €/tCO2`~ lag+ `Auction Volume tCO2`+ `Maximum Bid €/tCO2`+
#                                    oil_price,
 #                           family=gaussian, data=DE_only_with_lag)
#summary(DE_only_oil_fit)
#durbinWatsonTest(DE_only_oil_fit)

#pacf(DE_only_with_lag$oil_price, lag=20, pl=TRUE)

```

Fur Poland und nicht fur Deutschland:

```{r}
PL_only_auction<-ets_stock_and_oil_data |> filter(AuctionType=="PL") |> filter(
                                          is.na(`Auction Price €/tCO2`)==FALSE
)
PL_only_auction_price<- PL_only_auction$`Auction Price €/tCO2`
pacf(PL_only_auction_price)


#Auction Price Lag
Auction_Price_lag<- PL_only_auction$`Auction Price €/tCO2`[1:length(PL_only_auction$`Auction Price €/tCO2`)-1]
Auction_Price_lag<- c(0,Auction_Price_lag)
PL_only_with_lag<- PL_only_auction |> mutate(
                                          Auction_Price_lag= Auction_Price_lag, .after=`Auction Price €/tCO2`
)

#Maximum bid lag
Max_bid_lag<- PL_only_auction$`Maximum Bid €/tCO2`[1:length(PL_only_auction$`Maximum Bid €/tCO2`)-1]
Max_bid_lag<- c(0,Max_bid_lag)
PL_only_with_lag<- PL_only_with_lag |> mutate(
                                          Max_bid_lag= Max_bid_lag, .after=`Maximum Bid €/tCO2`
)

#Avg bids per bidPLr lag
Avg_num_bids_per_bidder_lag<- PL_only_auction$`Average number of bids per bidder`[1:length(PL_only_auction$`Average number of bids per bidder`)-1]
Avg_num_bids_per_bidder_lag<- c(0,Avg_num_bids_per_bidder_lag)
PL_only_with_lag<- PL_only_with_lag |> mutate(
                                          Avg_num_bids_per_bidder_lag= Avg_num_bids_per_bidder_lag, .after=`Average number of bids per bidder`
)

#Avg vol per bidder
Avg_vol_per_bidder_lag<- PL_only_auction$`Average volume bid per bidder`[1:length(PL_only_auction$`Average volume bid per bidder`)-1]
Avg_vol_per_bidder_lag<- c(0,Avg_vol_per_bidder_lag)
PL_only_with_lag<- PL_only_with_lag |> mutate(
                                          Avg_vol_per_bidder_lag= Avg_vol_per_bidder_lag, .after=`Average volume bid per bidder`
)

#oil price lag
PL_only_with_lag<- PL_only_with_lag |> rename(
                        oil_price = `Europe Brent Spot Price FOB (Dollars per Barrel)`
)
oil_price_lag<- PL_only_with_lag$oil_price[1:length(PL_only_with_lag$oil_price)-1]
oil_price_lag<- c(0,oil_price_lag)
PL_only_with_lag<- PL_only_with_lag |> mutate(
                                          oil_price_lag= oil_price_lag, .after= oil_price
)

#ftse price lag:
PL_only_with_lag<- PL_only_with_lag |> rename(
                        ftse_price = Price
)
ftse_price_lag<- PL_only_with_lag$ftse_price[1:length(PL_only_with_lag$ftse_price)-1]
ftse_price_lag<- c(0,ftse_price_lag)
PL_only_with_lag<- PL_only_with_lag |> mutate(
                                          ftse_price_lag= ftse_price_lag, .after= ftse_price
)

#Auction Volume tCO2 lag:

Auction_vol_lag<- PL_only_auction$`Auction Volume tCO2`[1:length(PL_only_auction$`Auction Volume tCO2`)-1]
Auction_vol_lag<- c(0,Auction_vol_lag)
PL_only_with_lag<- PL_only_with_lag |> mutate(
                                          Auction_vol_lag= Auction_vol_lag, .after=`Auction Volume tCO2`
)

#PL_only_oil_fit<- glm(`Auction Price €/tCO2`~ lag+ `Auction Volume tCO2`+ `Maximum Bid €/tCO2`+
 #                                   oil_price,
  #                          family=gaussian, data=PL_only_with_lag)
#summary(PL_only_oil_fit)
#durbinWatsonTest(PL_only_oil_fit)

#pacf(PL_only_with_lag$oil_price, lag=20, pl=TRUE)

#So for Poland, I don't need to worry about autocorrelation at all
```
What about EUAA:
```{r}
EUAA_only_auction<-ets_stock_and_oil_data |> filter(AuctionType=="EUAA") |> filter(
                                          is.na(`Auction Price €/tCO2`)==FALSE
)
EUAA_only_auction_price<- EUAA_only_auction$`Auction Price €/tCO2`
pacf(EUAA_only_auction_price)


#Auction Price Lag
Auction_Price_lag<- EUAA_only_auction$`Auction Price €/tCO2`[1:length(EUAA_only_auction$`Auction Price €/tCO2`)-1]
Auction_Price_lag<- c(0,Auction_Price_lag)
EUAA_only_with_lag<- EUAA_only_auction |> mutate(
                                          Auction_Price_lag= Auction_Price_lag, .after=`Auction Price €/tCO2`
)

#Maximum bid lag
Max_bid_lag<- EUAA_only_auction$`Maximum Bid €/tCO2`[1:length(EUAA_only_auction$`Maximum Bid €/tCO2`)-1]
Max_bid_lag<- c(0,Max_bid_lag)
EUAA_only_with_lag<- EUAA_only_with_lag |> mutate(
                                          Max_bid_lag= Max_bid_lag, .after=`Maximum Bid €/tCO2`
)

#Avg bids per bidPLr lag
Avg_num_bids_per_bidder_lag<- EUAA_only_auction$`Average number of bids per bidder`[1:length(EUAA_only_auction$`Average number of bids per bidder`)-1]
Avg_num_bids_per_bidder_lag<- c(0,Avg_num_bids_per_bidder_lag)
EUAA_only_with_lag<- EUAA_only_with_lag |> mutate(
                                          Avg_num_bids_per_bidder_lag= Avg_num_bids_per_bidder_lag, .after=`Average number of bids per bidder`
)

#Avg vol per bidder
Avg_vol_per_bidder_lag<- EUAA_only_auction$`Average volume bid per bidder`[1:length(EUAA_only_auction$`Average volume bid per bidder`)-1]
Avg_vol_per_bidder_lag<- c(0,Avg_vol_per_bidder_lag)
EUAA_only_with_lag<- EUAA_only_with_lag |> mutate(
                                          Avg_vol_per_bidder_lag= Avg_vol_per_bidder_lag, .after=`Average volume bid per bidder`
)

#oil price lag
EUAA_only_with_lag<- EUAA_only_with_lag |> rename(
                        oil_price = `Europe Brent Spot Price FOB (Dollars per Barrel)`
)
oil_price_lag<- EUAA_only_with_lag$oil_price[1:length(EUAA_only_with_lag$oil_price)-1]
oil_price_lag<- c(0,oil_price_lag)
EUAA_only_with_lag<- EUAA_only_with_lag |> mutate(
                                          oil_price_lag= oil_price_lag, .after= oil_price
)

#ftse price lag:
EUAA_only_with_lag<- EUAA_only_with_lag |> rename(
                        ftse_price = Price
)
ftse_price_lag<- EUAA_only_with_lag$ftse_price[1:length(EUAA_only_with_lag$ftse_price)-1]
ftse_price_lag<- c(0,ftse_price_lag)
EUAA_only_with_lag<- EUAA_only_with_lag |> mutate(
                                          ftse_price_lag= ftse_price_lag, .after= ftse_price
)

#Auction Volume tCO2 lag:

Auction_vol_lag<- EUAA_only_auction$`Auction Volume tCO2`[1:length(EUAA_only_auction$`Auction Volume tCO2`)-1]
Auction_vol_lag<- c(0,Auction_vol_lag)
EUAA_only_with_lag<- EUAA_only_with_lag |> mutate(
                                          Auction_vol_lag= Auction_vol_lag, .after=`Auction Volume tCO2`
)
```

```{r}

#Merging, beginning with EU lag
EU_lag<- EU_only_with_lag |> select('Date',
                                    'AuctionType', 
                                    'Auction_Price_lag',
                                    'Max_bid_lag',
                                    'Avg_num_bids_per_bidder_lag',
                                    'Avg_vol_per_bidder_lag',
                                    'Auction_vol_lag',
                                    'oil_price_lag',
                                    'ftse_price_lag'
                                    )
ets_EU_only_lag<- merge(x=ets_stock_and_oil_data, y= EU_lag, by=c('Date', 'AuctionType'), all.x=TRUE)
ets_EU_only_lag<- ets_EU_only_lag |> select(-c(Open:Change..))


lagColnamesVec<- c('Auction_Price_lag',
                                    'Max_bid_lag',
                                    'Avg_num_bids_per_bidder_lag',
                                    'Avg_vol_per_bidder_lag',
                                    'Auction_vol_lag',
                                    'oil_price_lag',
                                    'ftse_price_lag')

#DE lag:
DE_lag<- DE_only_with_lag |> select('Date',
                                    'AuctionType', 
                                    'Auction_Price_lag',
                                    'Max_bid_lag',
                                    'Avg_num_bids_per_bidder_lag',
                                    'Avg_vol_per_bidder_lag',
                                    'Auction_vol_lag',
                                    'oil_price_lag',
                                    'ftse_price_lag'
                                    )
ets_EU_DE_with_lag<- merge(x=ets_EU_only_lag, y=DE_lag, by=c('Date', 'AuctionType'), all.x=TRUE)


for(name in lagColnamesVec){
  x=paste0(name,".x")
  y=paste0(name,".y")
  ets_EU_DE_with_lag[,x]<-ifelse(is.na(ets_EU_DE_with_lag[,x]), 
                                  ets_EU_DE_with_lag[,y],
                                  ets_EU_DE_with_lag[,x])
}


ets_EU_DE_with_lag<- ets_EU_DE_with_lag |> rename(
                                    'Auction_Price_lag'= 'Auction_Price_lag.x',
                                    'Max_bid_lag'= 'Max_bid_lag.x',
                                    'Avg_num_bids_per_bidder_lag'='Avg_num_bids_per_bidder_lag.x',
                                    'Avg_vol_per_bidder_lag'='Avg_vol_per_bidder_lag.x',
                                    'Auction_vol_lag'= 'Auction_vol_lag.x',
                                    'oil_price_lag'= 'oil_price_lag.x',
                                    'ftse_price_lag'='ftse_price_lag.x'
                                    ) |> select(-c('Auction_Price_lag.y',
                                    'Max_bid_lag.y',
                                    'Avg_num_bids_per_bidder_lag.y',
                                    'Avg_vol_per_bidder_lag.y',
                                    'Auction_vol_lag.y',
                                    'oil_price_lag.y',
                                    'ftse_price_lag.y'))

#PL lag:
PL_lag<- PL_only_with_lag |> select('Date',
                                    'AuctionType', 
                                    'Auction_Price_lag',
                                    'Max_bid_lag',
                                    'Avg_num_bids_per_bidder_lag',
                                    'Avg_vol_per_bidder_lag',
                                    'Auction_vol_lag',
                                    'oil_price_lag',
                                    'ftse_price_lag'
                                    )
ets_EU_DE_PL_with_lag<- merge(x=ets_EU_DE_with_lag, 
                              y= PL_lag, by = c('Date', 'AuctionType'),
                              all.x=TRUE)

for(name in lagColnamesVec){
  x=paste0(name,".x")
  y=paste0(name,".y")
  ets_EU_DE_PL_with_lag[,x]<-ifelse(is.na(ets_EU_DE_PL_with_lag[,x]), 
                                  ets_EU_DE_PL_with_lag[,y],
                                  ets_EU_DE_PL_with_lag[,x])
}

ets_EU_DE_PL_with_lag<- ets_EU_DE_PL_with_lag |> rename(
                                    'Auction_Price_lag'= 'Auction_Price_lag.x',
                                    'Max_bid_lag'= 'Max_bid_lag.x',
                                    'Avg_num_bids_per_bidder_lag'='Avg_num_bids_per_bidder_lag.x',
                                    'Avg_vol_per_bidder_lag'='Avg_vol_per_bidder_lag.x',
                                    'Auction_vol_lag'= 'Auction_vol_lag.x',
                                    'oil_price_lag'= 'oil_price_lag.x',
                                    'ftse_price_lag'='ftse_price_lag.x'
                                    ) |> select(-c('Auction_Price_lag.y',
                                    'Max_bid_lag.y',
                                    'Avg_num_bids_per_bidder_lag.y',
                                    'Avg_vol_per_bidder_lag.y',
                                    'Auction_vol_lag.y',
                                    'oil_price_lag.y',
                                    'ftse_price_lag.y'))



#Adding EUAA lag:

EUAA_lag<- EUAA_only_with_lag |> select('Date',
                                    'AuctionType', 
                                    'Auction_Price_lag',
                                    'Max_bid_lag',
                                    'Avg_num_bids_per_bidder_lag',
                                    'Avg_vol_per_bidder_lag',
                                    'Auction_vol_lag',
                                    'oil_price_lag',
                                    'ftse_price_lag'
                                    )

ets_representative_lag<- merge(x=ets_EU_DE_PL_with_lag,
                               y=EUAA_lag, by = c('Date', 'AuctionType'),
                               all.x=TRUE)

for(name in lagColnamesVec){
  x=paste0(name,".x")
  y=paste0(name,".y")
  ets_representative_lag[,x]<-ifelse(is.na(ets_representative_lag[,x]), 
                                  ets_representative_lag[,y],
                                  ets_representative_lag[,x])
}

ets_representative_lag<- ets_representative_lag |> rename(
                                    'Auction_Price_lag'= 'Auction_Price_lag.x',
                                    'Max_bid_lag'= 'Max_bid_lag.x',
                                    'Avg_num_bids_per_bidder_lag'='Avg_num_bids_per_bidder_lag.x',
                                    'Avg_vol_per_bidder_lag'='Avg_vol_per_bidder_lag.x',
                                    'Auction_vol_lag'= 'Auction_vol_lag.x',
                                    'oil_price_lag'= 'oil_price_lag.x',
                                    'ftse_price_lag'='ftse_price_lag.x'
                                    ) |> select(-c('Auction_Price_lag.y',
                                    'Max_bid_lag.y',
                                    'Avg_num_bids_per_bidder_lag.y',
                                    'Avg_vol_per_bidder_lag.y',
                                    'Auction_vol_lag.y',
                                    'oil_price_lag.y',
                                    'ftse_price_lag.y'))

```

Checking Autocorrelation structures for various variables:
```{r}
ets_with_rep_lag<- na.omit(ets_representative_lag)
pacf(ets_with_rep_lag$`Maximum Bid €/tCO2`)
pacf(ets_with_rep_lag$`Average number of bids per bidder`)
pacf(ets_with_rep_lag$`Average volume bid per bidder`)
pacf(ets_with_rep_lag$oil_price)
pacf(ets_with_rep_lag$ftse_price)
pacf(ets_with_rep_lag$`Total Amount of Bids`)
pacf(ets_with_rep_lag$`Auction Volume tCO2`)
pacf(ets_with_rep_lag$`Auction Price €/tCO2`)
```

```{r}
#Lags have been created above. So let's run a first differences procedure.

#Starting by transforming the variables:

ets_first_diff<- ets_representative_lag |> rename(
                  oil_price= `Europe Brent Spot Price FOB (Dollars per Barrel)`,
                  ftse_price = Price
)

ets_modeling_data<- ets_first_diff

ets_first_diff<-ets_first_diff |> mutate(
                  Auction_Price_fd = `Auction Price €/tCO2`- Auction_Price_lag,
                  Max_bid_fd = `Maximum Bid €/tCO2` - Max_bid_lag,
                  Avg_bids_fd= `Average number of bids per bidder` - Avg_num_bids_per_bidder_lag,
                  Avg_vol_fd= `Average volume bid per bidder`- Avg_vol_per_bidder_lag,
                  Auction_vol_fd = `Auction Volume tCO2`- Auction_vol_lag,
                  oil_price_fd = oil_price - oil_price_lag,
                  ftse_price_fd = ftse_price -ftse_price_lag
) |> select (
            Auction_Price_fd,
            Max_bid_fd,
            Avg_bids_fd,
            Avg_vol_fd,
            Auction_vol_fd,
            oil_price_fd,
            ftse_price_fd
)


#Lets calculate a fit without lag and examine residuals:
fit_no_lag<- glm(`Auction Price €/tCO2` ~ `Maximum Bid €/tCO2`+
                                          `Average number of bids per bidder`+
                                          `Average volume bid per bidder`+
                                          `Auction Volume tCO2`+
                                            oil_price+ ftse_price, 
                 family=gaussian, data=ets_modeling_data)

res<- resid(fit_no_lag)
pacf(res,lag=10, pl=TRUE)


lag_transformed_fit<- glm(Auction_Price_fd ~ Max_bid_fd+Avg_bids_fd+Avg_vol_fd+
                            Auction_vol_fd+oil_price_fd+ftse_price_fd, 
                          family=gaussian(), data=ets_first_diff)

summary(lag_transformed_fit)

durbinWatsonTest(lag_transformed_fit)

#So this works, DW statistic is within good bounds. Now we need to run the regression without intercept, find the sample mean for each statistic and calculate the new fitted model. 

#Other models we use can run one of the other procedures to deal with time series correlations

```

So let's start with no pooling models. We have our complete pooling model above. Motivations for other models include the fact that we have four distinct auction types which might well vary. Further the lag above doesn't really account well for the differences between the models, especially the variation in their time series structure so no pooling and partial pooling models might well be very useful. 

1) No pooling for EU wide auctions:
```{r}
ets_EU_only<- ets_modeling_data |> filter(AuctionType=="EU")

#Let's start with a no lag fit, even though the DW for this is going to be terrible:

EU_fit_no_lag<- glm(`Auction Price €/tCO2` ~ `Maximum Bid €/tCO2`+
                                          `Average number of bids per bidder`+
                                          `Average volume bid per bidder`+
                                          `Auction Volume tCO2`+
                                            oil_price+ ftse_price, 
                 family=gaussian, data=ets_EU_only)

summary(EU_fit_no_lag)
durbinWatsonTest(EU_fit_no_lag)
res<- resid(fit_no_lag)
pacf(res,lag=10, pl=TRUE)

#As expected. Let's run it with the lag:


EU_ets_first_diff<- ets_EU_only|> mutate(
                  Auction_Price_fd = `Auction Price €/tCO2`- Auction_Price_lag,
                  Max_bid_fd = `Maximum Bid €/tCO2` - Max_bid_lag,
                  Avg_bids_fd= `Average number of bids per bidder` - Avg_num_bids_per_bidder_lag,
                  Avg_vol_fd= `Average volume bid per bidder`- Avg_vol_per_bidder_lag,
                  Auction_vol_fd = `Auction Volume tCO2`- Auction_vol_lag,
                  oil_price_fd = oil_price - oil_price_lag,
                  ftse_price_fd = ftse_price -ftse_price_lag
) |> select (
            Auction_Price_fd,
            Max_bid_fd,
            Avg_bids_fd,
            Avg_vol_fd,
            Auction_vol_fd,
            oil_price_fd,
            ftse_price_fd
)
EU_lag_transformed_fit<- glm(Auction_Price_fd ~ Max_bid_fd+Avg_bids_fd+Avg_vol_fd+
                            Auction_vol_fd+oil_price_fd+ftse_price_fd, 
                          family=gaussian(), data=EU_ets_first_diff)

summary(EU_lag_transformed_fit)

durbinWatsonTest(EU_lag_transformed_fit)

#Note, this isn't the final model, we still need to fit a no intercept model and then adjust the intercept with the means for each data thing. Lets check residuals anyway

res<- resid(EU_lag_transformed_fit)
plot(fitted(EU_lag_transformed_fit), res)

```
 No pooling for German Auctions:
 
```{r}
ets_DE_only<- ets_modeling_data |> filter(AuctionType=="DE")

#Let's start with a no lag fit, even though the DW for this is going to be terrible:

DE_fit_no_lag<- glm(`Auction Price €/tCO2` ~ `Maximum Bid €/tCO2`+
                                          `Average number of bids per bidder`+
                                          `Average volume bid per bidder`+
                                          `Auction Volume tCO2`+
                                            oil_price+ ftse_price, 
                 family=gaussian, data=ets_DE_only)

summary(DE_fit_no_lag)
durbinWatsonTest(DE_fit_no_lag)
res<- resid(fit_no_lag)
pacf(res,lag=10, pl=TRUE)

#As expected. Let's run it with the lag:


DE_ets_first_diff<- ets_DE_only|> mutate(
                  Auction_Price_fd = `Auction Price €/tCO2`- Auction_Price_lag,
                  Max_bid_fd = `Maximum Bid €/tCO2` - Max_bid_lag,
                  Avg_bids_fd= `Average number of bids per bidder` - Avg_num_bids_per_bidder_lag,
                  Avg_vol_fd= `Average volume bid per bidder`- Avg_vol_per_bidder_lag,
                  Auction_vol_fd = `Auction Volume tCO2`- Auction_vol_lag,
                  oil_price_fd = oil_price - oil_price_lag,
                  ftse_price_fd = ftse_price -ftse_price_lag
) |> select (
            Auction_Price_fd,
            Max_bid_fd,
            Avg_bids_fd,
            Avg_vol_fd,
            Auction_vol_fd,
            oil_price_fd,
            ftse_price_fd
)
DE_lag_transformed_fit<- glm(Auction_Price_fd ~ Max_bid_fd+Avg_bids_fd+Avg_vol_fd+
                            Auction_vol_fd+oil_price_fd+ftse_price_fd, 
                          family=gaussian(), data=DE_ets_first_diff)

summary(DE_lag_transformed_fit)

durbinWatsonTest(DE_lag_transformed_fit)

#Note, this isn't the final model, we still need to fit a no intercept model and then adjust the intercept with the means for each data thing. Lets check residuals anyway

res<- resid(DE_lag_transformed_fit)
plot(fitted(DE_lag_transformed_fit), res)

```

No Pooling model Fur Poland und Nicht fur Deutschland:
```{r}
ets_PL_only<- ets_modeling_data |> filter(AuctionType=="PL")

#Let's start with a no lag fit, even though the DW for this is going to be terrible:

PL_fit_no_lag<- glm(`Auction Price €/tCO2` ~ `Maximum Bid €/tCO2`+
                                          `Average number of bids per bidder`+
                                          `Average volume bid per bidder`+
                                          `Auction Volume tCO2`+
                                            oil_price+ ftse_price, 
                 family=gaussian, data=ets_PL_only)

summary(PL_fit_no_lag)
durbinWatsonTest(PL_fit_no_lag)
res<- resid(fit_no_lag)
pacf(res,lag=10, pl=TRUE)

#As expected. Let's run it with the lag:


PL_ets_first_diff<- ets_PL_only|> mutate(
                  Auction_Price_fd = `Auction Price €/tCO2`- Auction_Price_lag,
                  Max_bid_fd = `Maximum Bid €/tCO2` - Max_bid_lag,
                  Avg_bids_fd= `Average number of bids per bidder` - Avg_num_bids_per_bidder_lag,
                  Avg_vol_fd= `Average volume bid per bidder`- Avg_vol_per_bidder_lag,
                  Auction_vol_fd = `Auction Volume tCO2`- Auction_vol_lag,
                  oil_price_fd = oil_price - oil_price_lag,
                  ftse_price_fd = ftse_price -ftse_price_lag
) |> select (
            Auction_Price_fd,
            Max_bid_fd,
            Avg_bids_fd,
            Avg_vol_fd,
            Auction_vol_fd,
            oil_price_fd,
            ftse_price_fd
)
PL_lag_transformed_fit<- glm(Auction_Price_fd ~ Max_bid_fd+Avg_bids_fd+Avg_vol_fd+
                            Auction_vol_fd+oil_price_fd+ftse_price_fd, 
                          family=gaussian(), data=PL_ets_first_diff)

summary(PL_lag_transformed_fit)

durbinWatsonTest(PL_lag_transformed_fit)

#DW is much too high. We will have to test the lag permutations for this and sort it out

#Note, this isn't the final model, we still need to fit a no intercept model and then adjust the intercept with the means for each data thing. Lets check residuals anyway

res<- resid(PL_lag_transformed_fit)
plot(fitted(PL_lag_transformed_fit), res)
```

EUAA:

```{r}
ets_EUAA_only<- ets_modeling_data |> filter(AuctionType=="EUAA")

#Let's start with a no lag fit, even though the DW for this is going to be terrible:

EUAA_fit_no_lag<- glm(`Auction Price €/tCO2` ~ `Maximum Bid €/tCO2`+
                                          `Average number of bids per bidder`+
                                          `Average volume bid per bidder`+
                                          `Auction Volume tCO2`+
                                            oil_price+ ftse_price, 
                 family=gaussian, data=ets_EUAA_only)

summary(EUAA_fit_no_lag)
durbinWatsonTest(EUAA_fit_no_lag)
res<- resid(fit_no_lag)
pacf(res,lag=10, pl=TRUE)

#As expected. Let's run it with the lag:


EUAA_ets_first_diff<- ets_EUAA_only|> mutate(
                  Auction_Price_fd = `Auction Price €/tCO2`- Auction_Price_lag,
                  Max_bid_fd = `Maximum Bid €/tCO2` - Max_bid_lag,
                  Avg_bids_fd= `Average number of bids per bidder` - Avg_num_bids_per_bidder_lag,
                  Avg_vol_fd= `Average volume bid per bidder`- Avg_vol_per_bidder_lag,
                  Auction_vol_fd = `Auction Volume tCO2`- Auction_vol_lag,
                  oil_price_fd = oil_price - oil_price_lag,
                  ftse_price_fd = ftse_price -ftse_price_lag
) |> select (
            Auction_Price_fd,
            Max_bid_fd,
            Avg_bids_fd,
            Avg_vol_fd,
            Auction_vol_fd,
            oil_price_fd,
            ftse_price_fd
)
EUAA_lag_transformed_fit<- glm(Auction_Price_fd ~ Max_bid_fd+Avg_bids_fd+Avg_vol_fd+
                            Auction_vol_fd+oil_price_fd+ftse_price_fd, 
                          family=gaussian(), data=EUAA_ets_first_diff)

summary(EUAA_lag_transformed_fit)

durbinWatsonTest(EUAA_lag_transformed_fit)

#DW is much too high. But the no lag model is fine, so for EUAA, that might be the best model. 

#Note, this isn't the final model, we still need to fit a no intercept model and then adjust the intercept with the means for each data thing. Lets check residuals anyway

res<- resid(EUAA_lag_transformed_fit)
plot(fitted(EUAA_lag_transformed_fit), res)
```

#What about partial pooling?

We're going to try just varying intercept first even though it seems likely that varying the slope might be useful simply because some models need the lag transformation and others don't and that might affect slope.


